### 2019-07-29 reports/thesis/chapters/3_compoundPoisson.tex
Add citations to compound Poisson

The compound Poisson distribution is studied here because of the compound Poission-like behaviour from the detection of photons. It is defined by defining a latent variable $Y\sim\poisson(\lambda)$ with probability mass function (p.m.f.) $\prob(Y=y)=\euler^{-\lambda}\frac{\lambda^y}{y!}$ for $y=0,1,2,\cdots$, where $\lambda>0$ is the Poisson rate parameter. Let $U_i$ be some independent and identically distributed (i.i.d.) latent random variables with probability density function (p.d.f.) $p_U(u)$ for $i=1,2,3,\cdots$. Let $X$ be a compound Poisson random variable where
\begin{equation}
  X|Y = \sum_{i=1}^{Y}U_i \ .
  \label{eq:compoundPoisson_X|Y}
\end{equation}
The p.d.f.~of $X$ can be obtained by marginalising the joint p.d.f.
\begin{equation}
  p_X(x)=\sum_{y=0}^\infty p_{X|Y}(x|y)\prob(Y=y) \quad\text{for }x\geqslant 0
  \ .
\end{equation}
It should be noted that $X=0$ if and only if $Y=0$ with probability $\prob(X=0)=\euler^{-\lambda}$. Then $X$ has probability mass at zero and probability density for positive numbers which results in the p.d.f.
\begin{equation}
  p_X(x) = 
  \begin{cases}
    \delta(x) \euler^{-\lambda}  & \text{ for } x=0 \\ 
    \sum_{y=1}^\infty p_{X|Y}(x|y)\euler^{-\lambda}\frac{\lambda^y}{y!} \quad\text{for } & \text{ for } x>0
  \end{cases}
\end{equation}
where $\delta(x)$ is the Dirac delta function.

@@

Photons emitted from the x-ray tube undergo attenuation when propagating through the test sample. Assuming no beam hardening, some photons are either absorbed or scattered, making them undetectable. Scattered photons may be detected but it is very rare. Attenuated photon energy remains unaffected so attenuation decreases the parameter $\lambda$ and the amount it decreases by depends on the attenuation coefficient of the material and the amount of material the x-ray attenuates. The parameters of the random variable $U_i$ remains unchanged because the energy of each photon remains the same after attenuation, assuming no beam hardening.

When the photons interact with the scintillator in the detector, they are converted into visible light. The visible light photons are then detected and converted into a digital signal. A quantum counter set the digital signal to be linear with to the number of photons detected \citep{whiting2006properties}. Let $X$ be the grey value observed, then
\begin{equation}
X = bY + \epsilon
\end{equation}
where $\epsilon\sim\normal(a,\kappa)$, $b$ and $a$ are some constant and $\kappa$ is the variance of electronic noise. In a quantum counter, the mean and variance of the grey value are
\begin{equation}
\expectation\left[X\right] = b\lambda + a
\end{equation}
and
\begin{equation}
\variance\left[X\right] = b^2\lambda + \kappa
\end{equation}
respectively. By eliminating $\lambda$
\begin{equation}
\variance\left[X\right] = b\expectation\left[X\right]+\kappa-ab \ ,
\end{equation}
a linear relationship between the variance and expectation of the grey value is obtained.

An energy integrating detector records the grey value as linear to the energy detected \citep{whiting2006properties}. The grey value $X$ is
\begin{equation}
X|Y = \sum_{i=1}^Y U_i + \epsilon \ .
\end{equation}
This is the compound Poisson with Normal noise added to it. The scale factor $b$ is not included as this can be absorbed into $U$. Using the result that $\expectation\left[X\right]=\expectation\expectation\left[X|Y\right]$ and $\variance\left[X\right] = \variance\expectation\left[X|Y\right] + \expectation\variance\left[X|Y\right]$, the mean and variance of the grey value are
\begin{equation}
\expectation\left[X\right] = \lambda\expectation\left[U\right]+a
\end{equation}
and
\begin{equation}
\variance\left[X\right] = \lambda \expectation\left[U^2\right]+\kappa
\end{equation}
respectively. Eliminating $\lambda$ obtains
\begin{equation}
\variance\left[X\right] = \dfrac{\expectation\left[U^2\right]}{\expectation\left[U\right]} \expectation\left[X\right] + \kappa - a\dfrac{\expectation\left[U^2\right]}{\expectation\left[U\right]} \ .
\end{equation}
By assuming no beam hardening, $\expectation\left[U\right]$ and $\expectation\left[U^2\right]$ remains constant, thus there is a linear relationship between the variance and expectation of the grey value. There are other types of detection schemes \citep{whiting2006properties} but it shall be not be considered here.

@@

Returning back to the compound Poisson with no electronic noise $X|Y = \sum_{i=1}^{Y}U_i$. Let the moment generating function (m.g.f.) of $X$ be $M_X(\theta)=\expectation\left[\euler^{X\theta}\right]$. It can be shown that the m.g.f.~is
\begin{equation}
  M_X(\theta)=
  \exp\left[
    \lambda
    \left(
      M_U(\theta)-1
    \right)
  \right] \ .
\end{equation}
The derivation is shown in Appendix \ref{chapter:appendix_compoundPoissonMgf}. Moments of $X$ can be obtained from the m.g.f.~by differentiating it and setting it to zero. In other words, $\expectation[X^r]=M_X^{(r)}(\theta)$. Then it can be shown that
\begin{align}
  \expectation\left[X\right]&=\lambda\expectation\left[U\right]
  \\
  \variance\left[X\right] &= \lambda\expectation\left[U^2\right]
\end{align}
which agrees with the results in the previous section. The derivation is shown in Appendix \ref{chapter:appendix_compoundPoissonMgf}.

@@

It can be shown for large $\lambda$, the Normal approximation is
\begin{equation}
  X\sim\normal\left(\frac{\lambda\alpha}{\beta},\frac{\lambda\alpha(\alpha+1)}{\beta^2}\right) \ .
\end{equation}
The solution is shown in Appendix \ref{chapter:appendix_normalApproximation}.

@@

For the compound Poisson-gamma distribution, the saddle point approximation is given as 
\begin{multline}
  p_X(x)\approx
  \frac{\left(\lambda\alpha\beta^\alpha\right)^{\frac{1}{2(\alpha+1)}}\euler^{-\lambda}}{\sqrt{2\pi(\alpha+1)}}x^{-\frac{\alpha+2}{2(\alpha+1)}}
  \euler^{-x\beta}
  \exp\left[x^{\frac{\alpha}{\alpha+1}}
    \frac{(\lambda\beta^\alpha)^{\frac{1}{\alpha+1}}(\alpha+1)}{\alpha^{\frac{\alpha}{\alpha+1}}}
  \right]
  \\
  \text{for }x>0
  \label{eq:saddle_point_approx}
\end{multline}
with the derivation shown in Appendix \ref{chapter:appendix_saddlepoint}. The approximation is not well defined for $x=0$.

@@

The EM algorithm \citep{dempster1977maximum} is proposed as a method to estimate the parameters of a $X\sim\CPoisson(\lambda,\alpha,\beta)$ random variable given a sample of measurements of it $\left\{x_1,x_2,x_3,\cdots,x_n\right\}$.

>>>>

The compound Poisson distribution is studied here because of the compound Poission-like behaviour from the detection of photons \citep{whiting2002signal, elbakri2003efficient, whiting2006properties}. It is defined by defining a latent variable $Y\sim\poisson(\lambda)$ with probability mass function (p.m.f.) $\prob(Y=y)=\euler^{-\lambda}\frac{\lambda^y}{y!}$ for $y=0,1,2,\cdots$, where $\lambda>0$ is the Poisson rate parameter. Let $U_i$ be some independent and identically distributed (i.i.d.) latent random variables with probability density function (p.d.f.) $p_U(u)$ for $i=1,2,3,\cdots$. Let $X$ be a compound Poisson random variable where
\begin{equation}
  X|Y = \sum_{i=1}^{Y}U_i \ .
  \label{eq:compoundPoisson_X|Y}
\end{equation}
The p.d.f.~of $X$ can be obtained by marginalising the joint p.d.f.
\begin{equation}
  p_X(x)=\sum_{y=0}^\infty p_{X|Y}(x|y)\prob(Y=y) \quad\text{for }x\geqslant 0
  \ .
\end{equation}
It should be noted that $X=0$ if and only if $Y=0$ with probability $\prob(X=0)=\euler^{-\lambda}$. Then $X$ has probability mass at zero and probability density for positive numbers which results in the p.d.f.
\begin{equation}
  p_X(x) = 
  \begin{cases}
    \delta(x) \euler^{-\lambda}  & \text{ for } x=0 \\ 
    \sum_{y=1}^\infty p_{X|Y}(x|y)\euler^{-\lambda}\frac{\lambda^y}{y!} \quad\text{for } & \text{ for } x>0
  \end{cases}
\end{equation}
where $\delta(x)$ is the Dirac delta function.

@@

Photons emitted from the x-ray tube undergo attenuation when propagating through the test sample. Assuming no beam hardening, some photons are either absorbed or scattered, making them undetectable. Scattered photons may be detected but it is very rare \citep{cantatore2011introduction}. Attenuated photon energy remains unaffected so attenuation decreases the parameter $\lambda$ and the amount it decreases by depends on the attenuation coefficient of the material and the amount of material the x-ray attenuates. The parameters of the random variable $U_i$ remains unchanged because the energy of each photon remains the same after attenuation, assuming no beam hardening.

When the photons interact with the scintillator in the detector, they are converted into visible light. The visible light photons are then detected and converted into a digital signal. A quantum counter set the digital signal to be linear with to the number of photons detected \citep{whiting2006properties}. Let $X$ be the grey value observed, then
\begin{equation}
X = bY + \epsilon
\end{equation}
where $\epsilon\sim\normal(a,\kappa)$, $b$ and $a$ are some constant and $\kappa$ is the variance of electronic noise. In a quantum counter, the mean and variance of the grey value are
\begin{equation}
\expectation\left[X\right] = b\lambda + a
\end{equation}
and
\begin{equation}
\variance\left[X\right] = b^2\lambda + \kappa
\end{equation}
respectively. By eliminating $\lambda$
\begin{equation}
\variance\left[X\right] = b\expectation\left[X\right]+\kappa-ab \ ,
\end{equation}
a linear relationship between the variance and expectation of the grey value \citep{ma2012varaince} is obtained.

An energy integrating detector records the grey value as linear to the energy detected \citep{whiting2006properties}. The grey value $X$ is
\begin{equation}
X|Y = \sum_{i=1}^Y U_i + \epsilon \ .
\end{equation}
This is the compound Poisson with Normal noise added to it. The scale factor $b$ is not included as this can be absorbed into $U$. Using the result that $\expectation\left[X\right]=\expectation\expectation\left[X|Y\right]$ and $\variance\left[X\right] = \variance\expectation\left[X|Y\right] + \expectation\variance\left[X|Y\right]$, the mean and variance of the grey value are
\begin{equation}
\expectation\left[X\right] = \lambda\expectation\left[U\right]+a
\end{equation}
and
\begin{equation}
\variance\left[X\right] = \lambda \expectation\left[U^2\right]+\kappa
\end{equation}
respectively. Eliminating $\lambda$ obtains
\begin{equation}
\variance\left[X\right] = \dfrac{\expectation\left[U^2\right]}{\expectation\left[U\right]} \expectation\left[X\right] + \kappa - a\dfrac{\expectation\left[U^2\right]}{\expectation\left[U\right]} \ .
\end{equation}
By assuming no beam hardening, $\expectation\left[U\right]$ and $\expectation\left[U^2\right]$ remains constant, thus there is a linear relationship between the variance and expectation of the grey value \citep{yang2009evaluation}. There are other types of detection schemes \citep{whiting2006properties} but it shall be not be considered here.

@@

Returning back to the compound Poisson with no electronic noise $X|Y = \sum_{i=1}^{Y}U_i$. Let the moment generating function (m.g.f.) of $X$ be $M_X(\theta)=\expectation\left[\euler^{X\theta}\right]$. It can be shown that the m.g.f.~is
\begin{equation}
  M_X(\theta)=
  \exp\left[
    \lambda
    \left(
      M_U(\theta)-1
    \right)
  \right]
\end{equation}
\citep{gatto2010saddlepoint}. The derivation is shown in Appendix \ref{chapter:appendix_compoundPoissonMgf}. Moments of $X$ can be obtained from the m.g.f.~by differentiating it and setting it to zero. In other words, $\expectation[X^r]=M_X^{(r)}(\theta)$. Then it can be shown that
\begin{align}
  \expectation\left[X\right]&=\lambda\expectation\left[U\right]
  \\
  \variance\left[X\right] &= \lambda\expectation\left[U^2\right]
\end{align}
which agrees with the results in the previous section. The derivation is shown in Appendix \ref{chapter:appendix_compoundPoissonMgf}.

@@

It can be shown for large $\lambda$, the Normal approximation \citep{shevtsova2014on} is
\begin{equation}
  X\sim\normal\left(\frac{\lambda\alpha}{\beta},\frac{\lambda\alpha(\alpha+1)}{\beta^2}\right) \ .
\end{equation}
The solution is shown in Appendix \ref{chapter:appendix_normalApproximation}.

@@

For the compound Poisson-gamma distribution, the saddle point approximation \citep{jensen1991saddlepoint} is given as 
\begin{multline}
  p_X(x)\approx
  \frac{\left(\lambda\alpha\beta^\alpha\right)^{\frac{1}{2(\alpha+1)}}\euler^{-\lambda}}{\sqrt{2\pi(\alpha+1)}}x^{-\frac{\alpha+2}{2(\alpha+1)}}
  \euler^{-x\beta}
  \exp\left[x^{\frac{\alpha}{\alpha+1}}
    \frac{(\lambda\beta^\alpha)^{\frac{1}{\alpha+1}}(\alpha+1)}{\alpha^{\frac{\alpha}{\alpha+1}}}
  \right]
  \\
  \text{for }x>0
  \label{eq:saddle_point_approx}
\end{multline}
with the derivation shown in Appendix \ref{chapter:appendix_saddlepoint}. The approximation is not well defined for $x=0$.

@@

The EM algorithm \citep{dempster1977maximum} is proposed as a method to estimate the parameters of a $X\sim\CPoisson(\lambda,\alpha,\beta)$ random variable given a sample of measurements of it $\left\{x_1,x_2,x_3,\cdots,x_n\right\}$. The use of the EM algorithm here for the compound Poisson is similar to \cite{elbakri2003statistical, xu2009electronic}.
